---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)

library(tidyverse)

deaths <- av %>%
  pivot_longer(cols = starts_with("Death"),
               names_to = "Time",
               values_to = "Death") %>%
  mutate(Time = parse_number(Time)) %>%
  filter(Death != "")

head(deaths)

deaths_per_avenger <- deaths %>%
  group_by(Name.Alias) %>%
  summarise(total_deaths = n())

avg_deaths <- deaths_per_avenger %>%
  summarise(average_deaths = mean(total_deaths))

avg_deaths
```


Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 
```{r}
returns <- av %>%
  pivot_longer(
    cols = Return1:Return5,
    names_to = "Time",
    values_to = "Return"
  ) %>%
  mutate(Time = parse_number(Time))
returns
```
Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 

## Individually

For each team member, copy this part of the report. 

#Person 1
Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.
 "Given the Avengers’ 53 years in operation and overall mortality rate, fans of the comics can expect one current or former member to die every seven months or so, with a permanent death occurring once every 20 months."

### Include the code
Make sure to include the code to derive the (numeric) fact for the statement
```{r}
total_deaths <- deaths %>%
  filter(Death == "YES") %>%
  nrow()

years <- 53
months <- years * 12
months_per_death <- months / total_deaths

months_per_death
```


### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.
The data shows that the average time between deaths is 7.146067 months, which is very close to the claim of one death every seven months, so the statement is pretty much accurate.

#Person 2
Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement: There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death.

> Quote the statement you are planning to fact-check.

### Include the code
# join deaths and returns
```{r}
av2 <- av %>%
  mutate(id = row_number())
deaths2 <- av2 %>%
  pivot_longer(
    cols = Death1:Death5,
    names_to = "Time",
    values_to = "Death"
  ) %>%
  mutate(
    Time = parse_number(Time),
    Death = tolower(Death)
  )

returns2 <- av2 %>%
  pivot_longer(
    cols = Return1:Return5,
    names_to = "Time",
    values_to = "Return"
  ) %>%
  mutate(
    Time = parse_number(Time),
    Return = tolower(Return)
  )

death_return <- deaths2 %>%
  left_join(returns2, by = c("id", "Time"))

fact_check <- death_return %>%
  filter(Death == "yes", Time %in% c(1, 2, 3)) %>%
  mutate(group = case_when(
    Time == 1 ~ "first death",
    Time %in% c(2, 3) ~ "second or third death"
  )) %>%
  group_by(group) %>%
  summarise(
    total_deaths = n(),
    total_returns = sum(Return == "yes", na.rm = TRUE),
    recovery_rate = total_returns / total_deaths
  )

fact_check
```
Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Based on my analysis, the statement is not supported by my results. The recovery rate after a first death is 80%, while the recovery rate after a second or third death is 83.33%. This means the chance of returning after a second or third death is not lower than the chance of returning after a first death, so the statement is incorrect according to my calculation.

#Person 3
Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

#Person 4
Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.
